Tweet Classification: Filtering \\ of Twitter for Company-relevant Tweets

### 2012-12-02 paper/paper.tex
updated formatting and content of paper

\section{Method}
\subsection{Corpus creation}\\
By accessing the Twitter "Firehose" stream through their public API, we harvested 115,000 tweets using broad keyword filters with the goal of collecting a superset of relevant tweets. This provides us a smaller set of features in the overall corpus. The keywords used for the company of choice during this round of testing, Apple, were collected manually from information on the company's website and the Wikipedia page about Apple:\\
...\\
\\
This set of keywords clearly allows a large number of false positives:\\
nice fish and chip.. good pie apple :-) http://t.co/KVDN5r5l \\
20 chicken nuggets, chips, big mac, and a vanilla milkshake LIFE IS GOOD \\
1 more day whoop plus safari ride yay \\
shley staples an apple.  "AHHHH!!!!  Apple juice in my eye!" \\
I ain't seen uncle Mac in a while\\
\\
We define a tweet to be "about" Apple if the tweet mentions the company (via direct reference or stock symbol), one of its products, or a service it offers. We make the assumption that spam will be filtered from the testing set. This could be done via an existing Naive Bayes classifier, but we simply did not include instances of spam in our corpus during the manual labeling process. We collected our tweets over the course of twelve hours. On Twitter, a single message posted is often reposted, or "re-tweeted," many more times in a very short timespan. Because of the small timespan of our data collection, a large number of retweets (effectively duplicates) appeared in the dataset. We filtered these duplicates out of the dataset as well to prevent over-fitting in our model. After filtering, our original set of 115,000 tweets was reduced to 65,000. A representative sample of non-spam, non-duplicate tweets has composition of 60\% topical instances and 40\% false positives [need to analyze this for sure].
A computer-posted spam tweet: I've collected 10,650 gold coins! http://t.co/H0V0O6yP #ipad, #ipadgames, #gameinsight\\
A retweet, signified by "RT": RT @PHILerNotebook: Trying to fix my Mac. This gray loading bar takes 10 minutes to start up. Grr. Shall take it to an _Apple_ _store_ soon! \\
\\
We hand labeled 2000 tweets for use in training and testing, including only English-language tweets in our training set. \\
Data sets and scripts used for processing text data are available on the project website.

@@

We do all machine learning with the Mallet[2] library developed by the University of Massachusetts. Mallet includes abstractions for a data processing pipeline, during which we use both existing and custom processing "Pipes" to transform the data, beginning in our case with raw tweets text and ending in Mallet's internal FeatureVector format. We built several Pipes to analyze the effect of different filters and processes on the accuracy of classification:

@@

\subsubsection{Tokenization}
Twitter data, and user-generated web content in general, provides unique challenges to an automatic tokenizing system due to the relatively small number of features present and high level of noise in the feature set [citation and more details]. Features on twitter include excessive and irregular punctuation, frequent misspellings, emoticons, URLs, emoticons, and hashtags, to name a few.

Cool ranch , 4 berry sundae , apple juice #yessuh #latenight #snack http://t.co/ttc3vujp \\
I just poured apple juice on my cereal.. Gah. #SoTired \\
I hate Siri.. -___- \\

We approach this problem using a layered feature extraction system using regular expressions. Some features will often contain ambiguous or incorrect matches to other tokens. A prime example of this is the emoticon :/, which appears in every http link: http:/t.co/p9weubf [bold the emoticon somehow]. To alleviate such problems, we begin by tokenizing in order: URLs, emoticons, usernames, and hashtags. At this point, we are left with a fragmented tweet containing myriad punctuation, capitalization, and creative spellings. As done by Potts (2011) we normalize the length of all letter sequences greater than two, as in English, these are invariably due to user-added emphasis, and condensing the larger set of tokens focuses the feature set.\\

#news can't wait for my signed #OoRITE2OutNow it's emotional here right now loooool #OoRITE2OutNow https://t.co/r4n4cn6C \\

Capitalization\\
As presented in our results, because of irregular capitalization, two tokens likely to be semantically identical with respect to company classification may be represented in multiple cases: {Nooo, NOOO, nooo, NoOo}. To alleviate this, we attempt three different approaches. In the first, we simply lowercase all remaining text (note that many URLs become invalid after modifying case). In the second, we lowercase any text that is not in all caps. This serves to retain acronyms. In the third, we lowercase everything, leaving the case of all strings containing "apple" (case insensitive) alone. We made this decision based on the semantic difference between the capitalization of Apple the company and apple the fruit. This can be generalized to a larger set of keywords and applied to other companies in a similar fashion.

@@

\section{Contributions}

We have built several reusable Java components that extend the Mallet API for processig of data from Twitter. We created a TweetJsonIterator that accepts Twitter's JSON-formatted tweets and creates Mallet Instances containing their values for easy processing by Mallet's pipeline during model training. We have also created several Pipe implementations that can be reused or easily modified to suit a similar Twitter processing. These include FixEmoticons, Link2Title, Stemmer, SpellCheck, and TwitterFeatures. All code is available through our project website.

@@

\section{References}
\begin{thebibliography}{99}
@article{potts2011,
    author    = "Christopher Potts",
    title     = "Sentiment Symposium Tutorial: Tokenizing",
    journal   = "",
    year      = "2011"
}
@inproceedings{Pak10,
  added-at = {2010-06-01T10:23:44.000+0200},
  address = {Valletta, Malta},
  author = {Pak, Alexander and Paroubek, Patrick},
  biburl = {http://www.bibsonomy.org/bibtex/25656c3bb1adf00c58a85e3204096961c/frederik},
  booktitle = {Proceedings of the Seventh conference on International Language Resources and Evaluation (LREC'10)},
  date = {19-21},
  interhash = {ac930b0459a3c8a2fc2d74c52a475026},
  intrahash = {5656c3bb1adf00c58a85e3204096961c},
  isbn = {2-9517408-6-7},
  keywords = {imported},
  language = {english},
  month = may,
  publisher = {European Language Resources Association (ELRA)},
  timestamp = {2010-06-01T10:23:44.000+0200},
  title = {Twitter as a Corpus for Sentiment Analysis and Opinion Mining},
  url = {http://www.lrec-conf.org/proceedings/lrec2010/pdf/385_Paper.pdf},
  year = 2010
}
@inproceedings{Ruiz:2012:CFT:2124295.2124358,
 author = {Ruiz, Eduardo J. and Hristidis, Vagelis and Castillo, Carlos and Gionis, Aristides and Jaimes, Alejandro},
 title = {Correlating financial time series with micro-blogging activity},
 booktitle = {Proceedings of the fifth ACM international conference on Web search and data mining},
 series = {WSDM '12},
 year = {2012},
 isbn = {978-1-4503-0747-5},
 location = {Seattle, Washington, USA},
 pages = {513--522},
 numpages = {10},
 url = {http://doi.acm.org/10.1145/2124295.2124358},
 doi = {10.1145/2124295.2124358},
 acmid = {2124358},
 publisher = {ACM},
 address = {New York, NY, USA},
 keywords = {financial time series, micro-blogging, social networks},
} 
\end{thebibliography}

@@

[2] Mallett, University of Massachussetts Department of Computer Science\\

[3] Sanders Analytics (http://www.sananalytics.com/lab/twitter-sentiment/)

>>>>

\section{Method}
\subsection{Corpus creation}
By accessing the Twitter "Firehose" stream through their public API, we harvested 115,000 tweets using broad keyword filters with the goal of collecting a superset of relevant tweets. This provides us a smaller set of features in the overall corpus. The keywords used for the company of choice during this round of testing, Apple, were collected manually from information on the company's website and the Wikipedia page about Apple:\\
...\\

This set of keywords clearly allows a large number of false positives:\\
nice fish and chip.. good pie apple :-) http://t.co/KVDN5r5l \\
20 chicken nuggets, chips, big mac, and a vanilla milkshake LIFE IS GOOD \\
1 more day whoop plus safari ride yay \\
Ashley staples an apple.  "AHHHH!!!!  Apple juice in my eye!" \\
I ain't seen uncle Mac in a while\\

We define a tweet to be "about" Apple if the tweet mentions the company (via direct reference or stock symbol), one of its products, or a service it offers. We make the assumption that spam will be filtered from the testing set. This could be done via an existing Naive Bayes classifier, but we simply did not include instances of spam in our corpus during the manual labeling process. We collected our tweets over the course of twelve hours. On Twitter, a single message posted is often reposted, or "re-tweeted," many more times in a very short timespan. Because of the small timespan of our data collection, a large number of retweets (effectively duplicates) appeared in the dataset. We filtered these duplicates out of the dataset as well to prevent over-fitting in our model. After filtering, our original set of 115,000 tweets was reduced to 65,000. A representative sample of non-spam, non-duplicate tweets has composition of 60\% topical instances and 40\% false positives [need to analyze this for sure].
A computer-posted spam tweet: I've collected 10,650 gold coins! http://t.co/H0V0O6yP \#ipad, \#ipadgames, \#gameinsight\\
A retweet, signified by "RT": RT @PHILerNotebook: Trying to fix my Mac. This gray loading bar takes 10 minutes to start up. Grr. Shall take it to an \_Apple\_ \_store\_ soon! \\

We hand labeled 2000 tweets for use in training and testing, including only English-language tweets in our training set. \\
Data sets and scripts used for processing text data are available on the project website.\\

@@

We do all machine learning with the Mallet\cite{McCallumMALLET} library developed by the University of Massachusetts. Mallet includes abstractions for a data processing pipeline, during which we use both existing and custom processing "Pipes" to transform the data, beginning in our case with raw tweets text and ending in Mallet's internal FeatureVector format. We built several Pipes to analyze the effect of different filters and processes on the accuracy of classification. Not every Pipe proved to increase accuracy--some decreased it. We present the descriptions of and motivations for these Pipes here.

@@

\subsubsection{Tokenization}
Twitter data, and user-generated web content in general, provides unique challenges to an automatic tokenization system due to the high level of noise in the feature set [citation and more details]. Tweets are limited to 140 characters, which limits the total number of features possible. Some of the many natural language processing dilemmas present in tweets include excessive and irregular punctuation, frequent misspellings, emoticons, URLs, and hashtags, to name a few.

Cool ranch , 4 berry sundae , apple juice \#yessuh \#latenight \#snack http://t.co/ttc3vujp 
I just poured apple juice on my cereal.. Gah. \#SoTired 
I hate Siri.. -\_\_\_-

We approach this problem using a layered token extraction system. Using regular expressions, we begin by searching for tokens that may include other, more general tokens. A prime example of this is an emoticon appearing in a URL.

Searching for emoticons in this tweet will extract :/ as an emoticon, effectively breaking up the URL.

Now on my iphone ?? http://t.co/Rqr1RL2R

Additionally, users often include emoticons or links after text with no delimiting whitespace.

I listen to my iPod every day and it's just broke omg:(!!

Extracting tokens using this layered approach eliminates many of these problems.

After tokenizing these special features in order (URLs, emoticons, usernames, and hashtags) we are left with a fragmented tweet containing myriad punctuation, capitalization, and creative spellings. As done by Potts (2011) we normalize the length of all letter sequences greater than two, as in English, these are invariably due to user-added emphasis, and condensing the larger set of tokens focuses the feature set. In the following tweet, "loooool" is normalized to "loool".

\#news can't wait for my signed \#OoRITE2OutNow it's emotional here right now loooool \#OoRITE2OutNow https://t.co/r4n4cn6C

Note that such repetitions are often valid in URLs and some emoticons. Normalizing them prior to this step would make URLs invalid if using our link replacement Pipe.

As presented in our results, because of irregular capitalization, two tokens likely to be semantically identical with respect to company classification may be represented in multiple cases: {Nooo, NOOO, nooo, NoOo} is a prime example. To alleviate this, we attempt three different approaches. In the first, we simply lowercase all remaining tokens (note that many URLs become invalid after modifying case). In the second, we lowercase any text that is not in all caps. This serves to retain acronyms. In the third, we change all remaining tokens to lowercase, leaving the case of all strings containing "apple" (case insensitive) alone. We made this decision based on the semantic difference between the capitalization of Apple the company and apple the fruit. While correct capitalization is naturally not always present for this token, this is one case where capitalization serves to disambiguate. This same principle can be generalized to a larger set of keywords and applied to other companies in a similar fashion.

@@

\section{Contributions}
We have built several reusable Java components that extend the Mallet API for processing of data from Twitter. We created a TweetJsonIterator that accepts a file of Twitter's JSON-formatted tweets and creates training Instances containing their values for easy processing during model training. We have implemented several Pipes that can be reused or easily modified to suit a similar Twitter processing purpose. These include FixEmoticons, Link2Title, Stemmer, SpellCheck, and Tokenize, which serves as a more extensible tokenizer than Mallet's default. All code is available through our project website. We also provide a corpus of 115,000 tweets selected by Apple-related keyword and our labeled data set of 2000 tweets. The bash scripts used to filter the data for near-duplicates and spam are also available for download.

@@

\section{References}
\begin{thebibliography}{99}
@article{potts2011,
    author    = "Christopher Potts",
    title     = "Sentiment Symposium Tutorial: Tokenizing",
    journal   = "",
    year      = "2011"
}
@inproceedings{Pak10,
  added-at = {2010-06-01T10:23:44.000+0200},
  address = {Valletta, Malta},
  author = {Pak, Alexander and Paroubek, Patrick},
  biburl = {http://www.bibsonomy.org/bibtex/25656c3bb1adf00c58a85e3204096961c/frederik},
  booktitle = {Proceedings of the Seventh conference on International Language Resources and Evaluation (LREC'10)},
  date = {19-21},
  interhash = {ac930b0459a3c8a2fc2d74c52a475026},
  intrahash = {5656c3bb1adf00c58a85e3204096961c},
  isbn = {2-9517408-6-7},
  keywords = {imported},
  language = {english},
  month = may,
  publisher = {European Language Resources Association (ELRA)},
  timestamp = {2010-06-01T10:23:44.000+0200},
  title = {Twitter as a Corpus for Sentiment Analysis and Opinion Mining},
  url = {http://www.lrec-conf.org/proceedings/lrec2010/pdf/385\_Paper.pdf},
  year = 2010
}
@inproceedings{Ruiz:2012:CFT:2124295.2124358,
 author = {Ruiz, Eduardo J. and Hristidis, Vagelis and Castillo, Carlos and Gionis, Aristides and Jaimes, Alejandro},
 title = {Correlating financial time series with micro-blogging activity},
 booktitle = {Proceedings of the fifth ACM international conference on Web search and data mining},
 series = {WSDM '12},
 year = {2012},
 isbn = {978-1-4503-0747-5},
 location = {Seattle, Washington, USA},
 pages = {513--522},
 numpages = {10},
 url = {http://doi.acm.org/10.1145/2124295.2124358},
 doi = {10.1145/2124295.2124358},
 acmid = {2124358},
 publisher = {ACM},
 address = {New York, NY, USA},
 keywords = {financial time series, micro-blogging, social networks},
} 
@unpublished{McCallumMALLET,
    author = "Andrew Kachites McCallum",
    title = "MALLET: A Machine Learning for Language Toolkit",
    note = "http://mallet.cs.umass.edu",
    year = 2002,
}

\end{thebibliography}

@@
